---
title: "rqtl_pnut_bc"
author: "Ellen Risemberg"
date: "8/3/2021"
output: 
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# C3H x CC27 Backcross QTL Analysis 

See `rqtl_file_proc-pnutbc.R` for R/qtl file generation information. 

* A genotype: CC027
* B genotype: C3H/HeJ 

Genotypes are coded in R/qtl so that 1 = AA and 2 = AB. 

--------------------------------------------------------------------------------
### Setup
--------------------------------------------------------------------------------

```{r pkgs, message = FALSE}
library(qtl)
library(qtl2)
library(MASS)
library(MESS)
library(rstatix)
library(lme4qtl)
library(RColorBrewer)
library(onewaytests)
library(tidyverse)
library(ggpubr)
library(factoextra)
library(knitr)
source("code-dependencies/qtl_functions.R")
source("code-dependencies/plot_noX.R")
```

Load ggplot themes:

```{r ggthemes}
load_themes()
```

### Load data

```{r loadcross}
WB02 <- read.cross(format = "csv", file="derived_data/Rqtl_CC27xC3H_BC.csv",
                   na.strings=c("-", "NA", 'na'), genotypes=c("AA", "AB"),
                   alleles=c("A", "B"), crosstype=c("bc"))
WB02 <- jittermap(WB02)
summary(WB02)
raw_phenos <- WB02$pheno # save raw phenos 

WB02v2 <- convert2cross2(WB02) # convert to Rqtl2 cross object 
```

### Genetic map

```{r genmap}
plotMap(WB02)

estmapfile <- "derived_data/other_Robjects/estmap.Rdata"
if (file.exists(estmapfile)){
  load(estmapfile)
} else { # file doesn't exist 
  estmap <- est.map(WB02)
  save(estmap, file = estmapfile)
}

plotMap(WB02, estmap)
```

Save genetic map 

```{r savegenmap, include=F}
ensure_directory("figs/supplemental")
png("figs/supplemental/genetic_map.png", width = 650)
par(mar = c(5,5,4,1)+.1)
plotMap(WB02, main = "", ylab = "", xlab = "")
title(ylab="Location (cM)", cex.lab=1.8, line=2.5)
title(xlab="Chromosome", cex.lab=1.8)
title(main="Genetic map", cex.main=2)
dev.off()
```

Save genetic map compared to estimated map

```{r savegenmap2, include=F}
png("figs/supplemental/geneticmap_vs_estmap.png", width = 650)
par(mar = c(5,5,4,1)+.1)
plotMap(WB02, estmap, alternate.chrid = TRUE, ylab = "", xlab = "", main = "")
title(ylab="Location (cM)", cex.lab=1.8, line=2.5)
title(xlab="Chromosome", cex.lab=1.8)
dev.off()
```

### Genetic map statistics

largest gap:

```{r largestgap}
find_map_gaps(WB02v2$gmap, min_gap = 10)
```

Median physical distance in Mbp:

```{r distance}
alldistances <- vector()
  
for (i in 1:length(WB02v2$gmap)){
  positions <- WB02v2$gmap[[i]]
  distances <- positions[2:length(positions)] - positions[1:(length(positions)-1)]
  
  alldistances <- append(alldistances, distances)
}

median(alldistances)
```


## Exploratory data analysis  

Removed the following outliers from phenotype file prior to generation of R/qtl file:

* Cr_WB02_F0359 (Temp_0min is in the 300s) 
* Cr_WB02_F0480 (Temp_0min is 27.5 - much lower than rest of mice)

### Temperature histograms 

```{r}
par(mfrow = c(3,2))
hist(WB02$pheno$Temp_0min)
hist(WB02$pheno$Temp_15min)
hist(WB02$pheno$Temp_30min)
hist(WB02$pheno$Temp_45min)
hist(WB02$pheno$Temp_60min)
```

### Temperature trajectory plots

Set up data: 

```{r}
phenos = c('Temp_0min', 'Temp_15min', 'Temp_30min', 'Temp_45min','Temp_60min')  

## for ggplot
dat <- WB02$pheno$mouse_ID
for (i in 1:length(phenos)){
  y <- WB02$pheno[phenos[i]]
  dat <- cbind(dat, y)
}
dat <- as.data.frame(dat)
colnames(dat) <- c('mouse_ID', '0', '15', '30', '45', '60')
dat <- pivot_longer(dat, cols = c(2:6), names_to = 'Time', values_to = 'Temp')
```

Plot temperature trajectories for each mouse:

```{r}
bctemp <- ggplot(dat, aes(x = Time, y = Temp)) +
  geom_line(aes(group = mouse_ID), color = 'gray45', size=0.2) +
  labs(x = "Minutes post-challenge", y = "Body temperature") +
  scale_x_discrete(expand = c(0.03,0.03)) +
  scale_y_continuous(breaks = seq(from = 26, to = 40, by = 2)) +
  big_theme
bctemp

ensure_directory('figs/eda')
png('figs/eda/bc-temp.png', width = 550)
bctemp
dev.off()
```


### Hierarchical clustering:

Hierarchical clustering

```{r}
d <- dist(WB02$pheno[,phenos])
h <- hclust(d)
clusters <- cutree(h, k=3)
WB02$pheno[,'cluster'] <- clusters
```

Plot hierarchical clustering tree

```{r include=F}
#jpeg("figs/eda/traj_clusters.jpeg", width=1000)
plot(h, labels=FALSE, ylab = "", xlab = "", cex.main = 1.5)
title(ylab = "Height", cex.lab = 1.5, line=2.2)
title(xlab = "Distance", cex.lab = 1.5, line=0.5)
#dev.off()
```

Plot trajectories, colored by cluster

```{r}
dat$cluster <- rep(clusters, each=5)

pclust <- ggplot(dat, aes(x = Time, y = Temp)) +
  geom_line(aes(group = mouse_ID, color = as.factor(cluster)), size=0.5) + # change size to 0.2 
  scale_color_brewer(palette = 'Dark2', name = 'Cluster') +
  labs(x = "Minutes post-challenge", y = "Body temperature") +
  scale_x_discrete(expand = c(0.03,0.03)) +
  scale_y_continuous(breaks = seq(from = 26, to = 40, by = 2)) 
# + theme(legend.position = "none")

pclust

jpeg("figs/eda/trajectories-by-cluster.jpeg", width=600) # change width to 700
pclust + big_theme
dev.off()
```


### PCA

Principle components analysis based on observed phenotypes (temp and symptom score): 

```{r}
pca_data <- WB02$pheno[,c("Temp_0min", "Temp_15min", "Temp_30min", "Temp_45min", 
                          "Temp_60min", "Symptom_score")]
pca <- prcomp(pca_data)

pcaplot <- fviz_pca_ind(pca, habillage = WB02$pheno$cluster, geom="point", title = "", 
                        addEllipses = TRUE, mean.point = FALSE, legend.title = 'Cluster') 
pcaplot

jpeg("figs/eda/pca-by-cluster.jpeg", width = 700)
pcaplot + big_theme
dev.off()

pcaplot_pub <- fviz_pca_ind(pca, habillage = WB02$pheno$cluster, geom="point", title = "", 
                        addEllipses = TRUE, mean.point = FALSE, pointsize = 1, legend.title = 'Cluster') 
```

No major outliers 


### Summary statistics for temp trajectory 

Note that in calculating derived measures, the individual mouse's baseline temperature was used rather than the average of all mice's baseline temp. This is because from challenges done on CC strains, we know there is a significant strain effect on baseline temp. This indicates that there is a strain effect on baseline temp, outside of measurement error. Since backcross mice are outbred and not the same strain, we cannot assume they will all have the same baseline temp. (If variation in baseline temp were due to measurement error, we would want to use the average baseline temp to calculate AAC). 

See `temp_variation.R` for described analysis. 

**Derived measure #1: Area above the curve**

Based on time-series temperature data, plotted above. Calculated by subtracting the 
area under the curve from the total area of the rectangle formed by x = 60 min and 
y = Temp_0min[i] (the temp @ time 0 for the individual mouse). 

Illustrate concept for manuscript: 

```{r}
x <- c(0,15,30,45,60) # Minutes post-challenge (x-axis)
y1 <- dat$Temp[which(dat$mouse_ID == 'CR_WB02_F0005')]
y2 <- rep(y1[1], length(y1))

png("figs/supplemental/aac-illustration.png", width = 600)

plot(x, y1, type = "l", xlab = "", ylab = "")
title(xlab="Minutes post-exposure", cex.lab=1.5)
title(ylab="Body temperature", line=2.5, cex.lab=1.5)
legend(x=0, y=37.95, lty = c(1,2), legend = c('Temperature trajectory','Baseline temp'))
polygon(c(x, rev(x)), c(y2, rev(y1)), col = "#6BD7AF", density=10, angle=45)
lines(x, y1, lty=1)
abline(h = y1[1], lty=2)

dev.off()
```


```{r}
WB02 <- calc_auc(cross = WB02, steps = x, col.name = "Temp_aac", phenos = phenos)
plotPheno(WB02, pheno = 'Temp_aac')
```

**Derived measure #2: cluster**

Binarize cluster for more powerful mapping. Group 1 vs groups 2/3 will probably be the only one with statistical power, because group 3 is so small. 

Group 1 (no reaction) vs groups 2 and 3 (reaction):

```{r}
WB02$pheno$cluster_1vs23 <- ifelse(WB02$pheno$cluster == 1, 0, 1)
```

Ignoring group 1 (non-reactors), compare group 2 (recovery) vs group 3 (decline):

```{r}
WB02$pheno$cluster_2vs3 <- ifelse(WB02$pheno$cluster == 2, 0, ifelse(WB02$pheno$cluster == 3, 1, NA))
```

Group 1 or 2 (no reaction/recovery) vs group 3 (decline):

```{r}
WB02$pheno$cluster_12vs3 <- ifelse(WB02$pheno$cluster == 3, 1, 0)
```


### Correlation analysis

Plot symptom score vs temperature:

```{r}
ssvstemp <- ggplot(data = WB02$pheno, mapping = aes(x = as.factor(Symptom_score), y = Temp_aac)) + 
  geom_jitter(width = 0.15) + 
  labs(x = 'Symptom Score', y = 'Temperature - area above the curve') 

ssvstemp

jpeg('figs/supplemental/temp-vs-symptomscore.jpeg', width=600)
ssvstemp + big_theme
dev.off()
```

Relationship between symptom score and temperature: 

```{r}
summary(aov(Temp_aac ~ Symptom_score, data = WB02$pheno)) 
```


Plot temp vs IgE (don't expect a large correlation):

```{r}
igevstemp <- ggplot(data = WB02$pheno, mapping = aes(x = PNsIgE, y = Temp_aac)) + 
  geom_point() + 
  labs(x = 'Post-sensitization PNsIgE', y = 'Temperature - area above the curve')

igevstemp

jpeg('figs/supplemental/temp-vs-IgE.jpeg', width=600)
igevstemp + big_theme
dev.off()
```

Correlation between temp and PNsIgE:

```{r}
cor.test(WB02$pheno$Temp_aac, WB02$pheno$PNsIgE)
```


### Phenotype means and ranges 

In backcross mice: 

```{r}
print('Temp at 0 min')
mean(WB02$pheno$Temp_0min)
range(WB02$pheno$Temp_0min)
print('Temp at 15 min')
mean(WB02$pheno$Temp_15min)
range(WB02$pheno$Temp_15min)
print('Temp at 30 min')
mean(WB02$pheno$Temp_30min)
range(WB02$pheno$Temp_30min)
print('Temp at 45 min')
mean(WB02$pheno$Temp_45min)
range(WB02$pheno$Temp_45min)
print('Temp at 60 min')
mean(WB02$pheno$Temp_60min)
range(WB02$pheno$Temp_60min)

print('IgE')
mean(WB02$pheno$PNsIgE)
range(WB02$pheno$PNsIgE)

print('Symptom score')
mean(WB02$pheno$Symptom_score)
median(WB02$pheno$Symptom_score)
range(WB02$pheno$Symptom_score)
```


### Parent data

Load `parentdata` dataframe (with AAC / derived measures) and pparent plot  

```{r}
load('derived_data/other_Robjects/pparent.Rdata') # load pparent ggplot for Fig 1 
parentdata <- read_csv('derived_data/parentdata.csv')
```

Brown-Forsythe test - robust test for equal variances, even when underlying data are not normally distributed (so not using transformed data)

```{r}
# get auc metric for parent mice
parent_aac <- parentdata$Temp_aac 

# get auc metric for bc mice
bc_aac <- WB02$pheno$Temp_aac

bfdat <- data.frame(aac = as.numeric(c(parent_aac, bc_aac)), 
                    group = c(rep('parent', length(parent_aac)), rep('bc', length(bc_aac))))

bf.test(aac ~ group, bfdat)
```


### Figure 1

Histograms of symptom score and PNsIgE:

```{r sshist}
symptomhist <- ggplot(data = raw_phenos, aes(x = Symptom_score)) + 
  geom_histogram(binwidth = 0.5) + 
  labs(x = 'Symptom Score\n', y = 'Count') +
  pub_theme

symptomhist
```

```{r igehist}
igehist <- ggplot(data = raw_phenos, aes(x = PNsIgE)) + 
  geom_histogram(binwidth=2) + 
  labs(x = "Post-sensitization \nPNsIgE (\U00B5g/mL)", y = 'Count') + 
  pub_theme

igehist
```

Subpanels for trajectories of each cluster:

```{r}
dark2cols <- brewer.pal(n = 3, name = 'Dark2')

bctemp_c1 <- ggplot(dat[dat$cluster == 1,], aes(x = Time, y = Temp)) +
  geom_line(aes(group = mouse_ID), color = dark2cols[1], size=0.2) +
  labs(x = "Minutes post-challenge", y = "Body temperature") +
  scale_x_discrete(expand = c(0.03,0.03)) +
  scale_y_continuous(breaks = seq(from = 26, to = 40, by = 2), limits = c(25,40)) +
  pub_theme

bctemp_c2 <- ggplot(dat[dat$cluster == 2,], aes(x = Time, y = Temp)) +
  geom_line(aes(group = mouse_ID), color = dark2cols[2], size=0.2) +
  labs(x = "Minutes post-challenge", y = "Body temperature") +
  scale_x_discrete(expand = c(0.03,0.03)) +
  scale_y_continuous(breaks = seq(from = 26, to = 40, by = 2), limits = c(25,40)) +
  pub_theme

bctemp_c3 <- ggplot(dat[dat$cluster == 3,], aes(x = Time, y = Temp)) +
  geom_line(aes(group = mouse_ID), color = dark2cols[3], size=0.2) +
  labs(x = "Minutes post-challenge", y = "Body temperature") +
  scale_x_discrete(expand = c(0.03,0.03)) +
  scale_y_continuous(breaks = seq(from = 26, to = 40, by = 2), limits = c(25,40)) +
  pub_theme
```

Make Figure 1:

```{r warning=FALSE}
pparent_pub <- pparent + pub_theme
pclust_pub <- pclust + pub_theme
bctemp_pub <- bctemp + pub_theme
#pcaplot_pub <- pcaplot_pub + pub_theme

fig1 <- ggarrange(ggarrange(pparent_pub, pclust_pub, "", labels = c('A','B',''), 
                            ncol = 3, widths = c(0.5,0.47,0.03)), 
                  ggarrange(symptomhist, "", igehist, "", labels = c('C','','D',''), 
                            ncol = 4, widths = c(0.4,0.1,0.4,0.1)), 
                  nrow = 2)
fig1

ensure_directory('figs/fig1')
ggsave('figs/fig1/Figure1.jpeg', fig1)
```

Version with subpanels showing clusters separately:

```{r warning=FALSE}
fig1v2 <- ggarrange(ggarrange(pparent_pub, pclust_pub, "", labels = c('A','B',''), 
                              ncol = 3, widths = c(0.5,0.47,0.03)),
                    ggarrange("", bctemp_c1, bctemp_c2, bctemp_c3, labels = c('C'), 
                              ncol = 4, widths = c(0.03,0.3,0.3,0.3)),
                    ggarrange(symptomhist, "", igehist, "", labels = c('D','','E',''), 
                              ncol = 4, widths = c(0.4,0.1,0.4,0.1)), 
                    nrow = 3, heights = c(0.35, 0.3, 0.35))
fig1v2

ggsave('figs/fig1/Figure1wSubpanels.jpeg', fig1v2, width=2200, height=2200, units = 'px')
```


## Covariate analysis/transformation

Now that all summary statistics are calculated, import table with phenotype info:

```{r}
models <- read.csv("source_data/pnut_bc_models.csv")
```

Covariate analysis via ANOVA:

```{r}
for (i in 1:nrow(models)){
  mod.type <- models$type[i]
  
  if (mod.type == 'normal'){
    pheno <- models$colname[i]
    f <- paste(pheno, "~ Batch*Cage")
    df <- WB02$pheno
    assign("aov", do.call("aov", list(as.formula(f), data = as.name("df"))))
    print(paste("ANOVA results for ", pheno))
    print(anova(aov)) 
  }
}
```

**Results**: For all temperature phenotypes, there is *at most* a significant batch effect, i.e. there is no cage effect on any temp phenotypes. There are nominally significant batch and batch*cage interaction effects for the symptom score phenotype (both categorical and binarized), and there is a nominally significant cage effect on the diarrhea phenotype. 

### Transform data

Save raw phenotypes:

```{r}
raw_phenos <- WB02$pheno
```

Transform all data that we will model using the default (normal) model. 

* Temp_0min is normal enough 
* Temp_15min - TRINT 
* Temp_30min - TRINT
* Temp_45min - TRINT or standardized exp(x) - normality stats are v close
* Temp_60min - TRINT or standardized exp(x) - normality stats are v close 
* Temp_aac - TRINT
* Min_Temp - categorical (no transformation)
* Min_Temp_Time - categorical (no transformation)
* Symptom score - categorical (no transformation)
* Cluster - categorical / binary (no transformation)
* PNsIgE - square root 


```{r}
for (i in 2:nrow(models)){ # don't transform Temp_0min
  pheno <- models$colname[i]
  if (pheno == 'PNsIgE'){
    WB02$pheno[,pheno] <- sqrt(raw_phenos[,pheno])
  } else if(models$type[i] == 'normal'){
    WB02$pheno[,pheno] <- trint(raw_phenos[,pheno])
  } # else do nothing 
}
```


# Get residuals 

Instead of performing the following linear mixed model for each phenotype at each marker:
$$y_i = \mu + \beta_{geno}x_{geno,i} + bz_{b,i} + cz_{c,i} + \varepsilon_i$$
where $b$ and $c$ are random effects representing cage and batch; perform the following linear mixed model for each phenotype:
$$y_i = \mu + bz_{b,i} + cz_{c,i} + \varepsilon_i$$
get the residuals from the fit ("de-BLUP") and use `scanone` on the residuals to produce the model.

de-BLUP normal phenotypes (categorical/binary models do not assume normal residuals): 

```{r}
normal_phenos <- models$colname[which(models$type == 'normal')]
batch <- WB02$pheno[,'Batch']
cage <- WB02$pheno[,'Cage']

for (i in 1:length(normal_phenos)){
  colname <- normal_phenos[i]
  pheno <- WB02$pheno[,colname]
  formula <- paste0("pheno ~ ", ifelse(models$cov[i] == "", "", paste0(models$cov[i], " + ")), "(1|batch) + (1|batch:cage)")
  fit <- lmer(pheno ~ (1|batch) + (1|batch:cage))
  WB02$pheno[,colname] <- residuals(fit)
}
```


# Heritability Estimation 

Read genomic relationship matrix calculated with GCTA-GRM:

```{r}
grm <- readGRMBin(prefix = "source_data/heritability/plink-grm/pnut")

# create genomic relationship matrix from diagonal and off-diagonal elements 
G <- diag(grm$diag)
G[upper.tri(G)] <- grm$off
G[lower.tri(G)] <- t(G)[lower.tri(t(G))]
rownames(G) <- grm$id$V1
colnames(G) <- grm$id$V2
```

Estimate heritability:

```{r}
if(!file.exists('results/h2-from-bc-mice.csv')){ # if this code has not already run 
  set.seed(123)
  tbl1phenos <- c('Temp_0min', 'Temp_15min', 'Temp_30min', 'Temp_45min', 'Temp_60min',
                  'Min_Temp', 'Min_Temp_Time', 'PNsIgE', 'Symptom_score', 'cluster',
                  'Temp_aac')
  h2data <- data.frame(pheno = tbl1phenos,
                     h2 = rep(NA, length(tbl1phenos)), 
                     h2_lower = rep(NA, length(tbl1phenos)),
                     h2_upper = rep(NA, length(tbl1phenos))) ### add 2 p-value columns 

  # Manually estimate h2 for Temp_0min - not able to calculate confidence interval 
  # with code below (prof table has non-distinct values in .sigmaprop)
  mod1 <- relmatLmer(Temp_0min ~ (1|mouse_ID) + (1|Cage) + (1|Batch), data = WB02$pheno, 
                     relmat = list(mouse_ID = G))
  ci <- confint(mod1)
  h2data[1,2:4] <- c(VarProp(mod1)[1, 'prop'], ci[1,1], ci[1,2])
  
  for (i in 2:nrow(h2data)){
    pheno <- tbl1phenos[i]
  
    if (grepl('Temp', pheno) & (pheno != 'Temp_0min')){ # Temp phenotype, not baseline 
      formula1 <- paste0(pheno, " ~ Temp_0min + (1|Cage) + (1|Batch) + (1|mouse_ID)")
    } else {
      formula1 <- paste0(pheno, " ~ (1|mouse_ID) + (1|Cage) + (1|Batch)")
    }
    formula2 <- paste0(pheno, " ~ (1|mouse_ID)") # model with just polygenic term for CIs
  
    mod1 <- relmatLmer(formula1, data = WB02$pheno, relmat = list(mouse_ID = G))
    h2data[i,'h2'] <- VarProp(mod1)[1,'prop']
  
    mod2 <- relmatLmer(formula2, data = WB02$pheno, relmat = list(mouse_ID = G))
    prof <- profile(mod2)
    prof_prop <- varpropProf(prof)
    h2data[i,'h2_lower'] <- confint(prof_prop)[1,1]
    h2data[i,'h2_upper'] <- confint(prof_prop)[1,2]
  }

  write_csv(h2data, file = "results/h2-from-bc-mice.csv")
  
} else { # code has already run 
  h2data <- read_csv("results/h2-from-bc-mice.csv")
}
```



# Single-QTL Analysis 

Covariates are batch and cage, modeled as random effects. Sex is not included as a covariate bc all mice are female. Covariates are only allowed for the normal and binary models. Use Temp_0min as covariate on Temp_15min, Temp_30min, Temp_45min, and Temp_60min. 

## Models

Model choice is as follows based on distribution of phenotypes:

* Temp at all time points: normal
* Delta temp at all time points: normal
* Temperature aggregate/summary phenotypes: normal 
* Minimum temperature: normal 
* Time at minimum temperature: nonparametric (try POLR?)
* Symptom score: nonparametric (try POLR?)
* Symptom (binary): binary 
* Diarrhea (binary): binary

```{r}
kable(models)
```

First run `calc.genoprob()`:

```{r}
WB02 <- calc.genoprob(WB02)
```

Define covariate dataframe (include all phenotypes listed in `models$cov`, and `create_models` function will select only those listed for the phenotype):

```{r}
covar <- data.frame(Temp_0min = raw_phenos$Temp_0min)
```

Create models:

```{r warning=FALSE}
create_models(cross.obj = WB02, models = models, covar = covar, mod.dir = "derived_data/models/")
```



## Significance thresholds  

Permutation tests for LOD thresholds 

```{r}
create_perms(cross.obj = WB02, models = models, covar = covar, 
             perm.dir = "derived_data/perms/")
```


## Plot genome scans  

Plot genome scans with 5% and 10% significance thresholds shown by solid and dashed lines, respectively. 

```{r}
plot_scans(models = models)
```

Save genome scans 

```{r}
ensure_directory('figs/scans')
plot_scans(models = models[-which(models$obj == 'IgE'),], 
           save = TRUE, save.dir = 'figs/scans/', ylim = c(0, 4.5))
plot_scans(models = models[which(models$obj == 'IgE'),], 
           save = TRUE, save.dir = 'figs/scans/')
```


### Plot scans that will be at the bottom of each figure with modified X-axis 

Cluster scan with modified x-axis (bottom of Fig 2)

```{r}
png('figs/scans/cluster_1vs23_modX.png', width = 750)
plot_modX(cluster_1vs23, alternate.chrid = TRUE, xlab = "", ylab = "", main = "Temperature trajectory cluster", cex.main = 2, cex.axis = 2, bandcol = "gray90", ylim = c(0, 4.5))
title(ylab = "LOD", line = 2.5, cex.lab = 2)
title(xlab = "Chromosome", cex.lab = 2.3, line = 3.7) # cex.lab was 2
abline(h = summary(cluster_1vs23.p), lty=1:2)
dev.off()
```

PNsIgE scan with modified x-axis (only scan in Fig 4)

```{r}
png('figs/scans/PNsIgE_modX.png', width = 750)
plot_modX(IgE, alternate.chrid = TRUE, xlab = "", ylab = "", main = "Post-sensitization peanut-specific IgE (ug/mL)", cex.main = 2, cex.axis = 2, bandcol = "gray90")
title(ylab = "LOD", line = 2.5, cex.lab = 2)
title(xlab = "Chromosome", cex.lab = 2.3, line = 3.7) # cex.lab was 2
abline(h = summary(IgE.p), lty=1:2)
dev.off()
```

Symptom score scan with modified x-axis (bottom of Supp. Fig. 5)

```{r}
png('figs/supplemental/other-qtl/symptom_score_modX.png', width = 750)
plot_modX(sympScore, alternate.chrid = TRUE, xlab = "", ylab = "", main = "Symptom score", cex.main = 2, cex.axis = 2, bandcol = "gray90", ylim = c(0,4.5))
title(ylab = "LOD", line = 2.5, cex.lab = 2)
title(xlab = "Chromosome", cex.lab = 2.3, line = 3.7) # cex.lab was 2
abline(h = summary(SympScore.p), lty=1:2)
dev.off()
```


## Significant QTLs

Calculate 95% Bayes credible intervals for significant LOD peaks, and create a summary table with data for each LOD peak: marker, chromosome, position, LOD and positions of Bayes credible intervals.

```{r}
peaks <- doc_peaks(models)
print(peaks)
```

Create table for manuscript. Turn into tibble for easier processing:

```{r}
peakstbl <- as_tibble(peaks) %>% arrange(factor(chr, levels = c('10', '5', '17', '6', '18', '20', '9', '11')))
```

### Adjusted P-values

```{r table2, message=FALSE}
peakstbl$adj_pval <- rep(NA, nrow(peakstbl))

for (i in 1:nrow(peakstbl)){
  modname <- peakstbl$model[i]
  #permname <- models$perm.obj[which(models$obj == modname)]
  chr <- peakstbl$chr[i]
  
  gevname <- paste0(models$perm.obj[which(models$obj == modname)], '.gev')
  fitgev.sum <- summary.fevd(get(gevname))
  
  p <- pevd(q = as.numeric(peakstbl$lod[i]), 
            loc = fitgev.sum$par[1], 
            scale = fitgev.sum$par[2], 
            shape = fitgev.sum$par[3], 
            lower.tail = FALSE)
  # p <- mean(get(permname) > summary(get(modname))[as.integer(chr),'lod'])
  
  peakstbl$adj_pval[i] <- p
}

peakstbl$adj_pval <- trimws(format(peakstbl$adj_pval, digits=2))
peakstbl$adj_pval <- as.numeric(format(peakstbl$adj_pval, digits = 3))
```


### Genomic location 

Create `chr:position(interval)` column. First, split up bayes CI column into lower and upper limit. Then concatenate various data to produce column:

```{r}
peakstbl$BayesCIlower <- rep(NA, nrow(peakstbl))
peakstbl$BayesCIupper <- rep(NA, nrow(peakstbl))

for (i in 1:nrow(peakstbl)){
  peakstbl$BayesCIlower[i] <- strsplit(peakstbl$`Bayes CI`[i], split = ' - ')[[1]][1]
  peakstbl$BayesCIupper[i] <- strsplit(peakstbl$`Bayes CI`[i], split = ' - ')[[1]][2]
}

# Then concatenate  
peakstbl$chrint <- paste0(rep('Chr ',nrow(peakstbl)), 
                 peakstbl$chr, 
                 rep(': ',nrow(peakstbl)),
                 trimws(format(round(as.double(peakstbl$pos), 2), nsmall=2)),
                 rep(' (',nrow(peakstbl)),
                 trimws(format(round(as.double(peakstbl$BayesCIlower), 2), nsmall=2)),
                 rep('-',nrow(peakstbl)),
                 trimws(format(round(as.double(peakstbl$BayesCIupper), 2), nsmall=2)),
                 rep(')',nrow(peakstbl)))
```

Use phenotype name instead of model name: 

```{r}
peakstbl$Phenotype <- models$name[match(peakstbl$model, models$obj)]
```

### Unadjusted P-values 

```{r}
peakstbl$unadj_pval <- rep(NA, nrow(peakstbl))

for (i in 1:nrow(peakstbl)){
  marker = peakstbl$marker[i]
  mod = peakstbl$model[i]
  pheno.col = models$colname[models$obj == mod]
  covar <- models$cov[models$obj == mod]
  if (covar == "") {covar <- NULL}
  
  unadj.pval <- get_unadj_pval(WB02, pheno.col, marker, covar)
  peakstbl$unadj_pval[i] <- unadj.pval
}

peakstbl$unadj_pval <- format(peakstbl$unadj_pval, digits = 3)
```


### QTL names 

```{r}
peakstbl$QTL <- rep(NA, nrow(peakstbl))

qtl_pheno_counts <- peakstbl %>% group_by(chr) %>% dplyr::summarize(n = n())

peakstbl$QTL <- c(rep('Qpa1', qtl_pheno_counts$n[qtl_pheno_counts$chr==10]),
                  rep('Qpa2', qtl_pheno_counts$n[qtl_pheno_counts$chr==5]), 
                  rep('Qpa3', qtl_pheno_counts$n[qtl_pheno_counts$chr==17]), 
                  rep('Qpa4', qtl_pheno_counts$n[qtl_pheno_counts$chr==6]),
                  rep('Qpa5', qtl_pheno_counts$n[qtl_pheno_counts$chr==18]), 
                  rep('Qpa6', qtl_pheno_counts$n[qtl_pheno_counts$chr==20]), 
                  rep('Qpa7', qtl_pheno_counts$n[qtl_pheno_counts$chr==9]),
                  rep('Qpa8', qtl_pheno_counts$n[qtl_pheno_counts$chr==11]))
```


### Phenotypic variance explained

AKA heritability due to a QTL. Using formula (described in Rqtl guide pg 122)

$$h^2 = \frac{\text{var}\{E(y|g\}}{\text{var}\{y\}}$$

For backcross: 

$$h^2 = \frac{a^2}{a^2 + 4\sigma^2}$$

where $a = \mu_{AB} - \mu_{AA}$ and $\sigma^2$ is the residual variance. 

```{r}
peakstbl$phenotypic.var.expl <- rep(NA, nrow(peakstbl))

for (i in 1:nrow(peakstbl)){
  marker = peakstbl$marker[i]
  mod = peakstbl$model[i]
  pheno.col = models$colname[models$obj == mod]
  covar <- models$cov[models$obj == mod]
  if (covar == "") {covar <- NULL}
  
  h2 <- get_var_expl(WB02, pheno.col, marker, covar)
  
  peakstbl$phenotypic.var.expl[i] <- h2
}

peakstbl$phenotypic.var.expl <- paste0(as.numeric(format(peakstbl$phenotypic.var.expl, digits=3))*100, '%')
```

### Format and save 

Remove redundant data, re-order: 

```{r}
peakstbl <- peakstbl %>% 
  arrange(match(chr, c('10', '5', '17', '6', '18', '20', '9', '11')), as.numeric(adj_pval))

# now that it's sorted by adjusted p-value, add */**/*** (makes it a string so can't sort)
peakstbl$adj_pval <- paste(peakstbl$adj_pval, ifelse(peakstbl$adj_pval < 0.005, '(***)', ifelse(peakstbl$adj_pval < 0.05, '(**)', ifelse(peakstbl$adj_pval < 0.10, '(*)', NA))))

col_order <- c('QTL', 'Phenotype', 'chrint', 'unadj_pval', 'adj_pval', 'phenotypic.var.expl')
peakstbl <- peakstbl[,col_order]
```

Save to csv: 

```{r}
write.csv(peakstbl, file = "results/QTLsummary.csv", row.names = FALSE)
```


### Variance explained by Qpa1 and Qpa2 

Of temp-aac (most holistic measure of response)

```{r var}
WB02i <- fill.geno(WB02)
qpa1geno <- pull.geno(WB02i, chr = 10)[,'S2T100225813'] # chr10 marker most significantly associated with aac
qpa2geno <- pull.geno(WB02i, chr = 5)[,'gUNC9726629'] # chr5 marker most significantly associated with aac
pheno <- WB02i$pheno$Temp_aac

lmod <- lm(pheno ~ WB02i$pheno$Temp_0min + qpa1geno + qpa2geno)
summary(lmod)

# heritability calculation
sigma2 <- summary(lmod)$sigma^2
uAA1 <- mean(pheno[which(qpa1geno == 1)]) # 1 = AA
uAB1 <- mean(pheno[which(qpa1geno == 2)]) # 2 = AB 
a1 <- uAB1 - uAA1 
uAA2 <- mean(pheno[which(qpa2geno == 1)]) # 1 = AA
uAB2 <- mean(pheno[which(qpa2geno == 2)]) # 2 = AB 
a2 <- uAB2 - uAA2 

h2 <- (a1^2 + a2^2) / (a1^2 + a2^2 + 4*sigma2)
print(h2)
```


### Variance explained by Qpa1-3

Of temp-aac (most holistic measure of response)

```{r}
WB02i <- fill.geno(WB02)
qpa1geno <- pull.geno(WB02i, chr = 10)[,'S2T100225813'] # chr10 marker most significantly associated with aac
qpa2geno <- pull.geno(WB02i, chr = 5)[,'gUNC9726629'] # chr5 marker most significantly associated with aac
qpa3geno <- pull.geno(WB02i, chr = 17)[,'gUNC27787813'] # chr17 marker most significantly associated with T30 (not associated with aac)
pheno <- WB02i$pheno$Temp_aac

lmod <- lm(pheno ~ WB02i$pheno$Temp_0min + qpa1geno + qpa2geno + qpa3geno)
summary(lmod)

# heritability calculation
sigma2 <- summary(lmod)$sigma^2
uAA1 <- mean(pheno[which(qpa1geno == 1)]) # 1 = AA
uAB1 <- mean(pheno[which(qpa1geno == 2)]) # 2 = AB 
a1 <- uAB1 - uAA1 
uAA2 <- mean(pheno[which(qpa2geno == 1)]) # 1 = AA
uAB2 <- mean(pheno[which(qpa2geno == 2)]) # 2 = AB 
a2 <- uAB2 - uAA2 
uAA3 <- mean(pheno[which(qpa3geno == 1)]) # 1 = AA
uAB3 <- mean(pheno[which(qpa3geno == 2)]) # 2 = AB 
a3 <- uAB3 - uAA3 
h2 <- (a1^2 + a2^2 + a3^2) / (a1^2 + a2^2 +a3^2 + 4*sigma2)
print(h2)
```


### Variance of PNsIgE explained by Qpa7 and Qpa8 

```{r}
qpa7geno <- pull.geno(WB02, chr = 9)[,'mbackupUNC090355700'] 
qpa8geno <- pull.geno(WB02, chr = 11)[,'S3C111805292'] 
pheno <- WB02$pheno$PNsIgE

lmod <- lm(pheno ~ qpa7geno*qpa8geno)
summary(lmod)

# heritability calculation
sigma2 <- summary(lmod)$sigma^2
uAA1 <- mean(pheno[which(qpa7geno == 1)]) # 1 = AA
uAB1 <- mean(pheno[which(qpa7geno == 2)]) # 2 = AB 
a1 <- uAB1 - uAA1 
uAA2 <- mean(pheno[which(qpa8geno == 1)]) # 1 = AA
uAB2 <- mean(pheno[which(qpa8geno == 2)]) # 2 = AB 
a2 <- uAB2 - uAA2 
h2 <- (a1^2 + a2^2) / (a1^2 + a2^2 + 4*sigma2)
print(h2)
```


### Phenotype x genotype plots 

Plot PxG plots for all significant LOD peaks

```{r}
qtl.map <- data.frame(chr = c(10, 5, 17, 6, 18, 'X', 9, 11), qtl_name = paste0('Qpa', seq(1,8)))
plot_pxg(cross.obj = WB02, raw.data = raw_phenos, geno.map = list(A = "CC027", B = "C3H"),
         qtl.map = qtl.map, peaks = peaks, plot.type = 'pxg', theme = rmd_theme)
```

Save PxG plots 

```{r}
ensure_directory('figs/pxg')
plot_pxg(cross.obj = WB02, raw.data = raw_phenos, geno.map = list(A = "CC027", B = "C3H"),
         qtl.map = qtl.map, peaks = peaks, plot.type = 'pxg', theme = big_pxg_theme,
         save = TRUE, save.dir = 'figs/pxg/')

ensure_directory('figs/pxg/same_ylim')
plot_pxg(cross.obj = WB02, raw.data = raw_phenos, geno.map = list(A = "CC027", B = "C3H"),
         qtl.map = qtl.map, peaks = peaks, plot.type = 'pxg', theme = big_pxg_theme, ylim = c(27,40), 
         save = TRUE, save.dir = 'figs/pxg/same_ylim/')
```


#### Stacked bar plots isntead of scatter plots for categorical phenotypes 

Qpa2-cluster 

```{r barplot}
ggdat <- data.frame(Genotype = pull.geno(WB02i, 5)[,'UNC9309125'], 
                    Cluster = WB02$pheno$cluster)
ggdat$Cluster <- ifelse(ggdat$Cluster == 1, "Non-reactor", "Reactor")
ggdat$Cluster <- factor(ggdat$Cluster)
ggdat$Genotype <- ifelse(ggdat$Genotype == 1, 'CC027/CC027', 'CC027/C3H')
ggdat$Genotype <- factor(ggdat$Genotype, levels = c('CC027/CC027', 'CC027/C3H'))
ggdatsum <- ggdat %>% 
  dplyr::group_by(Genotype, Cluster) %>% 
  dplyr::summarize(n = n()) %>%
  mutate(Frequency = n / sum(n))

ensure_directory("figs/fig2")
png("figs/fig2/Qpa2-cluster.png", width = 700)
ggplot(data = ggdatsum, mapping = aes(x = Genotype, y = Frequency, fill = Cluster)) + 
  geom_bar(stat = 'identity', position = 'fill') +
  scale_fill_brewer(palette = 'Dark2') + 
  labs(x = bquote(italic(Qpa2)~(chr5)~Genotype)) +
  big_pxg_theme
dev.off()
```


Qpa4-time of min temp (categorical)

```{r barplot2}
ggdat <- data.frame(Genotype = pull.geno(WB02, chr=6)[,'mbackupUNC060210105'], 
                    Min_Temp_Time = WB02$pheno$Min_Temp_Time)
ggdat$Min_Temp_Time <- factor(ggdat$Min_Temp_Time)
ggdat$Genotype <- ifelse(ggdat$Genotype == 1, 'CC027/CC027', 'CC027/C3H')
ggdat$Genotype <- factor(ggdat$Genotype, levels = c('CC027/CC027', 'CC027/C3H'))
ggdatsum <- ggdat %>% 
  dplyr::group_by(Genotype, Min_Temp_Time) %>% 
  dplyr::summarize(n = n()) %>%
  mutate(Frequency = n / sum(n))

ensure_directory("figs/supplemental/other-qtl")
png("figs/supplemental/other-qtl/Qpa4-minTempTime.png", width = 700)
ggplot(data = ggdatsum, mapping = aes(x = Genotype, y = Frequency, fill = Min_Temp_Time)) + 
  geom_bar(stat = 'identity', position = 'fill') +
  scale_fill_brewer(palette = 'Dark2') + 
  labs(x = bquote(italic(Qpa4)~(chr6)~Genotype), fill = "Min temp time") + 
  big_pxg_theme
dev.off()
```


Qpa3-symptom score (categorical)

```{r barplot3}
ggdat2 <- data.frame(Genotype = pull.geno(WB02, chr=17)[,'S6Y171422147'], 'Symptom score' = WB02$pheno$Symptom_score)
ggdat2$Symptom.score <- factor(ggdat2$Symptom.score)
ggdat2$Genotype <- ifelse(ggdat2$Genotype == 1, 'CC027/CC027', 'CC027/C3H')
ggdat2$Genotype <- factor(ggdat2$Genotype, levels = c('CC027/CC027', 'CC027/C3H'))
ggdat2sum <- ggdat2 %>% 
  dplyr::group_by(Genotype, Symptom.score) %>% 
  dplyr::summarize(n = n()) %>%
  mutate(Frequency = n / sum(n))

png("figs/supplemental/other-qtl/Qpa3-sympScore.png", width = 720)
ggplot(data = ggdat2sum, mapping = aes(x = Genotype, y = Frequency, fill = Symptom.score)) + 
  geom_bar(stat = 'identity', position = 'fill') +
  scale_fill_brewer(palette = 'Dark2') + 
  labs(x = bquote(italic(Qpa3)~(chr17)~Genotype), fill = "Symptom score") +
  big_pxg_theme
dev.off()
```


### Extra PNsIgE peak analysis 

Confirm that chr9 and chr11 QTL (associated with PNsIgE) are not associated with any response phenotype:

```{r pnsige}
chr9m <- 'mbackupUNC090355700'

resp_phenos <- c('Temp_15min', 'Temp_30min', 'Temp_45min', 'Temp_60min',
                 'Min_Temp', 'Min_Temp_Time', 'Temp_aac', 'Symptom_score')
chr9m_lods <- c(T15[chr9m,'lod'], T30[chr9m,'lod'], T45[chr9m,'lod'], 
                T60[chr9m, 'lod'], minTemp[chr9m, 'lod'], minTempTime[chr9m, 'lod'], 
                sympScore[chr9m, 'lod'], Taac[chr9m, 'lod'])

chr9m_assoc <- data.frame(phenos = resp_phenos, 
                          lods = chr9m_lods, 
                          unadj_pval = rep(NA, length(chr9m_lods)))


for (i in 1:length(resp_phenos)){
  if (resp_phenos[i] == 'Symptom_score'){
    chr9m_assoc$unadj_pval[i] <- get_unadj_pval(WB02, pheno.col = resp_phenos[i],
                                           marker = chr9m)
  } else {
    chr9m_assoc$unadj_pval[i] <- get_unadj_pval(WB02, pheno.col = resp_phenos[i],
                                           marker = chr9m, covar = 'Temp_0min')
  }
}
```


```{r pnsige2}
chr11m <- 'S3C111805292'

chr11m_lods <- c(T15[chr11m,'lod'], T30[chr11m,'lod'], T45[chr11m,'lod'], 
                T60[chr11m, 'lod'], minTemp[chr11m, 'lod'], 
                minTempTime[chr11m, 'lod'], 
                sympScore[chr11m, 'lod'], Taac[chr11m, 'lod'])

chr11m_assoc <- data.frame(phenos = resp_phenos, 
                          lods = chr11m_lods, 
                          unadj_pval = rep(NA, length(chr11m_lods)))


for (i in 1:length(resp_phenos)){
  if (resp_phenos[i] == 'Symptom_score'){
    chr11m_assoc$unadj_pval[i] <- get_unadj_pval(WB02, pheno.col = resp_phenos[i],
                                           marker = chr11m)
  } else {
    chr11m_assoc$unadj_pval[i] <- get_unadj_pval(WB02, pheno.col = resp_phenos[i],
                                           marker = chr11m, covar = 'Temp_0min')
  }
}
```


# Multiple QTL Analysis 

marker genotypes from temp scans: 

```{r}
m5 <- 'UNC9309125' # chr5 marker from cluster scan
m10 <- 'UNC17654311' # chr10 marker from temp @ 15 min scan 
m17 <- 'S6Y171422147' # chr17 marker from temp @ 45 min scan 
mX <- 'SXX204357509' # chrX marker from temp @ 45 min scan 
m18 <- 'gUNC29817480' # chr18 marker from min temp scan 
m6 <- 'mbackupUNC060210105' # chr6 marker from time of min temp scan 

m5_genos <- pull.geno(WB02, 5)[,m5] 
m10_genos <- pull.geno(WB02, 10)[,m10] 
m17_genos <- pull.geno(WB02, 17)[,m17]
mX_genos <- pull.geno(WB02,'X')[,mX]
m18_genos <- pull.geno(WB02, 18)[,m18]
m6_genos <- pull.geno(WB02, 6)[,m6]
```

**Variable selection**

Via stepwise ANOVA

```{r stepwise}
temp_aac <- pull.pheno(WB02, pheno.col = 'Temp_aac')
n <- length(temp_aac)
t0 <- pull.pheno(WB02, pheno.col = 'Temp_0min')

full_mod <- lm(temp_aac ~ t0 + m5_genos + m10_genos + m17_genos + mX_genos + m18_genos + m6_genos)
```

Stepwise regression 

```{r stepwise2}
step_mod <- stepAIC(full_mod, direction = "both", trace = TRUE, k = log(n))
```

Is Qpa1*Qpa2 interaction significant? 

```{r}
mod1 <- lm(temp_aac ~ t0 + m10_genos + m5_genos + m17_genos + mX_genos + m18_genos + m6_genos)
mod6 <- lm(temp_aac ~ t0 + m10_genos + m5_genos + m17_genos + mX_genos + m18_genos + m6_genos + m10_genos*m5_genos)
anova(mod6, mod1)
```

*No*, interaction is not significant. 

So all QTL except for the one on chr6 is our best model (with AAC as the outcome). 

Does PNsIgE improve this model?

```{r}
ige_pheno <- pull.pheno(WB02, pheno.col = 'PNsIgE')

modwIgE <- lm(temp_aac ~ t0 + m5_genos + m10_genos + m17_genos + mX_genos + m18_genos + ige_pheno)
anova(modwIgE, step_mod)
```

*Yes.* 

Do PNsIgE QTL improve model?

Qpa7:

```{r}
qpa7geno <- pull.geno(WB02, chr = 9)[,'mbackupUNC090355700']
modwQpa7 <- lm(temp_aac ~ t0 + m10_genos + m5_genos + m17_genos + mX_genos + m18_genos + ige_pheno + qpa7geno)
anova(modwIgE, modwQpa7)
```

Qpa8: 

```{r}
qpa8geno <- pull.geno(WB02, chr = 11)[,'S3C111805292']
modwQpa8 <- lm(temp_aac ~ t0 + m10_genos + m5_genos + m17_genos + mX_genos + m18_genos + ige_pheno + qpa8geno)
anova(modwIgE, modwQpa8)
```

*No.*


The chr6 QTL (Qpa4) is associated with time of minimum temp. Are we sure the info captured by that phenotype is also captured in AAC (i.e. is AAC a representative phenotype for time of minimum temp)?

```{r}
outcome_phenos <- c('Temp_15min', 'Temp_30min', 'Temp_45min', 'Temp_60min', 'Min_Temp', 'Min_Temp_Time', 'cluster', 'Temp_aac', 'Symptom_score')
outcome_pca <- prcomp(WB02$pheno[,outcome_phenos], center = TRUE, scale = TRUE)


png('figs/supplemental/outcome-pca.png')
fviz_pca_var(outcome_pca, repel = TRUE)
dev.off()
#ggsave(filename = 'figs/supplemental/outcome-pca.jpeg')

png('figs/vars/scree-plot.png', width = 600)
fviz_screeplot(outcome_pca)
dev.off()
```

Min_Temp_Time appears to be capturing different information, so we probably still believe Qpa4. 

Run PCA without Min_Temp_Time:

```{r}
outcome_phenos.mtt <- c('Temp_15min', 'Temp_30min', 'Temp_45min', 'Temp_60min', 'Min_Temp', 'cluster', 'Temp_aac', 'Symptom_score')
outcome_pca.mtt <- prcomp(WB02$pheno[,outcome_phenos.mtt], center = TRUE, scale = TRUE)

png('figs/supplemental/outcome-pca-mtt.png')
fviz_pca_var(outcome_pca.mtt, repel = TRUE)
dev.off()

png('figs/vars/screeplot-mtt.png')
fviz_screeplot(outcome_pca.mtt)
dev.off()
```

Try factor analysis:

```{r}
ggsave('figs/vars/scree-plot.jpeg')
```

Clear elbow at 1, so try with 1 factor:

```{r}
temp_phenos <- c('Temp_15min', 'Temp_30min', 'Temp_45min', 'Temp_60min', 'Min_Temp', 'Min_Temp_Time', 'cluster', 'Temp_aac')
fa1 <- factanal(WB02$pheno[,temp_phenos], factors = 1)
fa1
```

Min_Temp_Time and Temp_0min have high uniqueness and are not strongly associated with Factor 1. 

Taken together, we believe that Min_Temp_Time is capturing different info from Temp-AAC, so we still believe Qpa4. 

### Do Qpa1 (chr10) and Qpa2 (chr5) interact or effect phenotype additively?

**Temp @ 15min** 

Plot temp at 15 min vs *Qpa1* genotype, colored by *Qpa2* genotype:

```{r}
pcolbym <- pxg_colbym(cross = WB02, pheno = raw_phenos$Temp_15min, marker = 'UNC17654311',
           marker2 = 'UNC9309125', geno.map = list(A = "CC027", B = "C3H"), 
           xlab = bquote(italic(Qpa1)~Genotype), ylab = 'Temp at 15 min', title = "", type = 'boxplot',
           theme = poster_theme)

pcolbym <- pcolbym + labs(fill = bquote(italic(Qpa2)~Genotype)) # use with type = 'boxplot'
#pcolbym <- pcolbym + labs(col = bquote(italic(Qpa2)~Genotype)) # use with type = 'scatter'

pcolbym

jpeg('figs/multi-qtl/T15-Qpa1-Qpa2-boxplot.jpeg', width=650)
pcolbym
dev.off()
```

T-test to compare Qpa2:CC027/CC027 mice to Qpa2:CC027/C3H mice among Qpa1:CC027/CC027 mice: 

```{r}
temp_15min <- WB02$pheno$Temp_15min
temp_15min_Qpa1hom <- temp_15min[which(m10_genos==1)] # homozygotes at Qpa1
m5_genos_Qpa1hom <- m5_genos[which(m10_genos==1)]
t.test(temp_15min_Qpa1hom ~ m5_genos_Qpa1hom)
```

T-test to compare Qpa1:CC027/CC027 mice to Qpa1:CC027/C3H mice among Qpa2:CC027/CC027 mice: 

```{r}
temp_15min_Qpa2hom <- temp_15min[which(m5_genos==1)] # homozygotes at Qpa2
m10_genos_Qpa2hom <- m10_genos[which(m5_genos==1)]
t.test(temp_15min_Qpa2hom ~ m10_genos_Qpa2hom)
```

T-test to compare mice that are homozygous at one allele and het at the other (should be the same): 

```{r}
temp_15min_Qpa1homQpa2het <- temp_15min[which((m10_genos == 1) & (m5_genos == 2))]
temp_15min_Qpa1hetQpa2hom <- temp_15min[which((m10_genos == 2) & (m5_genos == 1))]
t.test(temp_15min_Qpa1homQpa2het, temp_15min_Qpa1hetQpa2hom)
```

T-test to compare Qpa1:CC027/CC027 mice to Qpa1:CC027/C3H mice among Qpa2:CC027/C3H mice: 

```{r}
temp_15min_Qpa2het <- temp_15min[which(m5_genos==2)] # hets at Qpa2
m10_genos_Qpa2het <- m10_genos[which(m5_genos==2)]
t.test(temp_15min_Qpa2het ~ m10_genos_Qpa2het)
```

T-test to compare Qpa2:CC027/CC027 mice to Qpa2:CC027/C3H mice among Qpa1:CC027/C3H mice: 

```{r}
temp_15min_Qpa1het <- temp_15min[which(m10_genos==2)] # hets at Qpa1
m5_genos_Qpa1het <- m5_genos[which(m10_genos==2)]
t.test(temp_15min_Qpa1het ~ m5_genos_Qpa1het)
```


**Temp - AAC**

Plot temp AAC vs *Qpa1* genotype, colored by *Qpa2* genotype:

```{r}
pcolbymaac <- pxg_colbym(cross = WB02, pheno = WB02$pheno$Temp_aac, marker = 'UNC17654311',
           marker2 = 'UNC9309125', geno.map = list(A = "CC027", B = "C3H"), 
           xlab = bquote(italic(Qpa1)~Genotype), ylab = 'Temp - AAC (normalized)', title = "", 
           type = 'boxplot', theme = poster_theme)

pcolbymaac <- pcolbymaac + 
  scale_color_brewer(palette = 'Dark2', name = bquote(italic(Qpa2)~Genotype))
#pcolbymaac <- pcolbymaac + labs(col = bquote(italic(Qpa2)~Genotype)) # use with type = 'scatter'

pcolbymaac 

jpeg('figs/multi-qtl/Taac-Qpa1-Qpa2-boxplot.jpeg', width=650)
pcolbymaac
dev.off()
```

T-test to compare Qpa2:CC027/CC027 mice to Qpa2:CC027/C3H mice among Qpa1:CC027/CC027 mice: 

```{r}
temp_aac <- WB02$pheno$Temp_aac
temp_aac_Qpa1hom <- temp_aac[which(m10_genos==1)] # AAC pheno for homozygotes at Qpa1
m5_genos_Qpa1hom <- m5_genos[which(m10_genos==1)] # Qpa2 genotype for homozygotes at Qpa1 
t.test(temp_aac_Qpa1hom ~ m5_genos_Qpa1hom) 
```

T-test to compare Qpa1:CC027/CC027 mice to Qpa1:CC027/C3H mice among Qpa2:CC027/CC027 mice: 

```{r}
temp_aac_Qpa2hom <- temp_aac[which(m5_genos==1)] # AAC pheno for homozygotes at Qpa2
m10_genos_Qpa2hom <- m10_genos[which(m5_genos==1)] # Qpa1 genos for homozygotes at Qpa2 
t.test(temp_aac_Qpa2hom ~ m10_genos_Qpa2hom)
```

T-test to compare mice that are homozygous at one allele and het at the other (should be the same): 

```{r}
temp_aac_Qpa1homQpa2het <- temp_aac[which((m10_genos == 1) & (m5_genos == 2))]
temp_aac_Qpa1hetQpa2hom <- temp_aac[which((m10_genos == 2) & (m5_genos == 1))]
t.test(temp_aac_Qpa1homQpa2het, temp_aac_Qpa1hetQpa2hom)
```

T-test to compare Qpa1:CC027/CC027 mice to Qpa1:CC027/C3H mice among Qpa2:CC027/C3H mice: 

```{r}
temp_aac_Qpa2het <- temp_aac[which(m5_genos==2)] # AAC pheno for hets at Qpa2
m10_genos_Qpa2het <- m10_genos[which(m5_genos==2)] # Qpa1 geno for hets at Qpa2 
t.test(temp_aac_Qpa2het ~ m10_genos_Qpa2het)
```

T-test to compare Qpa2:CC027/CC027 mice to Qpa2:CC027/C3H mice among Qpa1:CC027/C3H mice: 

```{r}
temp_aac_Qpa1het <- temp_aac[which(m10_genos==2)] # AAC pheno for hets at Qpa1
m5_genos_Qpa1het <- m5_genos[which(m10_genos==2)] # Qpa2 geno for hets at Qpa1 
t.test(temp_aac_Qpa1het ~ m5_genos_Qpa1het)
```


Boxplot with significance bars:

```{r}
marker.genos <- WB02i$geno[[10]]$data[,'UNC17654311']
marker2.genos <- WB02i$geno[[5]]$data[,'UNC9309125']
  
df <- data.frame(geno = marker.genos,
                 geno2 = marker2.genos,
                 pheno = WB02$pheno$Temp_aac)

geno.labels = c("CC027/CC027", "CC027/C3H")
  
# map genotypes to genotype labels                 
df$geno <- mapvalues(df$geno, from = c(1,2), to = geno.labels)
df$geno2 <- mapvalues(df$geno2, from = c(1,2), to = geno.labels)
  
df$geno <- factor(df$geno, levels = geno.labels)
df$geno2 <- factor(df$geno2, levels = geno.labels)

jpeg("figs/multi-qtl/Qpa1Qpa2wsig.jpeg", width = 650)
ggplot(data = df, mapping = aes(x = geno, y = pheno, fill = geno2)) + 
     geom_boxplot() +
     scale_fill_brewer(palette = 'Dark2', name = bquote(italic(Qpa2)~Genotype)) + 
     geom_signif(annotation = "*", y_position = 2.5, xmin = 0.813, xmax = 1.185, tip_length = 0, textsize = 9) +
     geom_signif(annotation = "*", y_position = 3.2, xmin = 0.813, xmax = 1.81, tip_length = 0, textsize = 9) +
     geom_signif(annotation = "*", y_position = 1.8, xmin = 1.81, xmax = 2.188, tip_length = 0, textsize = 9) +
     geom_signif(annotation = "*", y_position = 2.8, xmin = 1.185, xmax = 2.188, tip_length = 0, textsize = 9) +
     labs(y = "Temp AAC (normalized)", x = bquote(italic(Qpa1)~Genotype)) +
     poster_theme
dev.off()


jpeg("figs/multi-qtl/Qpa1Qpa2wsig_v2.jpeg", width = 650)
ggplot(data = df, mapping = aes(x = geno, y = pheno, fill = geno2)) + 
     geom_boxplot() +
     scale_fill_brewer(palette = 'Dark2', name = bquote(italic(Qpa2)~Genotype)) + 
     geom_signif(annotation = "*", y_position = 2.5, xmin = 0.813, xmax = 1.185, tip_length = 0.01, textsize = 7) +
     geom_signif(annotation = "*", y_position = 3.1, xmin = 0.813, xmax = 1.81, tip_length = 0.01, textsize = 7) +
     geom_signif(annotation = "*", y_position = 1.8, xmin = 1.81, xmax = 2.188, tip_length = 0.01, textsize = 7) +
     geom_signif(annotation = "*", y_position = 2.7, xmin = 1.185, xmax = 2.188, tip_length = 0.01, textsize = 7) +
     labs(y = "Temp AAC (normalized)", x = bquote(italic(Qpa1)~Genotype)) +
     poster_theme
dev.off()


jpeg("figs/multi-qtl/Qpa1Qpa2.jpeg", width = 650, height = 450)
ggplot(data = df, mapping = aes(x = geno, y = pheno, fill = geno2)) + 
     geom_boxplot() +
     scale_fill_brewer(palette = 'Dark2', name = bquote(italic(Qpa2)~Genotype)) + 
     labs(y = "Temp AAC", x = bquote(italic(Qpa1)~Genotype)) +
     poster_theme
dev.off()
```


## Regress on number of CC027 alleles 

Calculate number of CC027 alleles: 

Coded as 1 = CC027/CC027 and 2 = CC027/C3H 
Recode m5 and m10 as 1 = CC027/C3H and 2 = CC027/CC027 (= number of CC027 alleles) 
and sum 

```{r}
# Qpa1
m10_numCC027 <- vector(length = length(m10_genos))
m10_numCC027[m10_genos == 1] <- 2
m10_numCC027[m10_genos == 2] <- 1

# Qpa2
m5_numCC027 <- vector(length = length(m5_genos))
m5_numCC027[m5_genos == 1] <- 2
m5_numCC027[m5_genos == 2] <- 1

numCC027 <- m10_numCC027 + m5_numCC027
```

Regress on number of CC027 alleles:

```{r}
fit <- lm(WB02$pheno$Temp_aac ~ numCC027)
summary(fit)
```


## Chr 5/10 haplotype vs cluster 

Plot chr 5/10 haplotype vs cluster

```{r}
m5m10df <- data.frame(m5.geno = m5_genos, 
                      m10.geno = m10_genos, 
                      cluster = pull.pheno(WB02, 'cluster'))

datasum <- m5m10df %>% 
  group_by(m5.geno, m10.geno, cluster) %>% 
  dplyr::summarise(n = n())

m5m10df$m5m10 <- paste0(m5m10df$m5.geno, "-", m5m10df$m10.geno)

jpeg("figs/fig3/m5m10cluster.jpeg", width=600)
ggplot(data = m5m10df, mapping = aes(x = m5m10, y = cluster)) + 
  geom_jitter(width = 0.2, height = 0.2) +
  scale_y_discrete(limits = c("1", "2", "3")) +
  scale_x_discrete(labels = c("1-1" = "Chr5:CC027/CC027\nChr10:CC027/CC027",
                              "1-2" = "Chr5:CC027/CC027\nChr10:CC027/C3H",
                              "2-1" = "Chr5:CC027/C3H\nChr10:CC027/CC027",
                              "2-2" = "Chr5:CC027/C3H\nChr10:CC027/C3H")) +
  xlab("Genotype") +
  rmd_theme
dev.off()
```



### Contingency table test 

Two-way contingency table: overall genotype vs cluster 

```{r}
datasum2way <- m5m10df %>% 
  group_by(m5m10, cluster) %>% 
  dplyr::summarise(n = n())

# 2-way table, poisson model 
mod2p <- glm(n ~ m5m10*cluster, datasum2way, family = 'poisson')

```

Three-way contingency table: chr 5 genotype, chr 10 genotype, cluster 

`datasum` is our contingency table

```{r}
modu <- glm(n ~ (m5.geno + m10.geno + cluster)^2, datasum, family=poisson)

mod
```


What's going on at chr17 (MHC locus)?

```{r}
m1 <- 'UNC17654311' # chr10
m2 <- 'gUNC27920115' # chr17
m1.genos <- WB02$geno[[10]]$data[,m1]
m2.genos <- WB02$geno[[17]]$data[,m2]
pheno = raw_phenos[,'Temp_30min']
geno.labels <- c('AA', 'AB')

df <- data.frame(chr10 = m1.genos, chr17 = m2.genos, pheno = pheno)

df$chr10 <- mapvalues(df$chr10, from = c(1,2), to = geno.labels)
df$chr17 <- mapvalues(df$chr17, from = c(1,2), to = geno.labels)
df$geno <- paste('chr10:', df$chr10, ' chr17:', df$chr17, sep="")

dfsum <- df %>% 
  group_by(geno) %>% 
  dplyr::summarise(mean = mean(pheno), sd = sd(pheno))

#jpeg("/Users/ellenrisemberg/Desktop/temp_30min_chr10AndChr17.jpeg", width = 700)
ggplot(data = df, mapping = aes(x = geno, y = pheno)) +
  geom_jitter(width = 0.1, height = 0) + 
  geom_errorbar(data = dfsum,
                mapping = aes(x = geno, y = mean, ymin = mean-sd, ymax = mean+sd),
                width = 0.3) +
  scale_x_discrete(labels = function(x) stringr::str_wrap(x, width = 7)) +
  labs(title = 'Phenotype vs genotype at chr10 and chr17 markers', 
       y = 'Temp at 30 min', x = 'Genotype')
#dev.off()
```


